\propertitle

### 2019-12-22 maxNt_v6.tex
update

\pdfoutput=1
%% Author: PGL  Porta Mana
%% Created: 2015-05-01T20:53:34+0200
%% Last-Updated: 2019-12-22T00:23:26+0100
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newif\ifarxiv
\arxivfalse
\ifarxiv\pdfmapfile{+classico.map}\fi
\newif\ifafour
\afourfalse% true = A4, false = A5
\newif\iftypodisclaim % typographical disclaim on the side
\typodisclaimtrue
\newcommand*{\memfontfamily}{zplx}
\newcommand*{\memfontpack}{newpxtext}
\documentclass[\ifafour a4paper,12pt,\else a5paper,10pt,\fi%extrafontsizes,%
onecolumn,oneside,article,%french,italian,german,swedish,latin,
british%
]{memoir}
\newcommand*{\updated}{\today}
\newcommand*{\firstdraft}{4 November 2015}
\newcommand*{\firstpublished}{***}
\newcommand*{\propertitle}{Inferring the total activity of a large neuronal population\\ from a small sample}
\newcommand*{\pdftitle}{\propertitle}
\newcommand*{\headtitle}{Inferences about large neuronal populations}
\newcommand*{\pdfauthor}{P.G.L.  Porta Mana, Y. Roudi, V. Rostami, E. Torre}
\newcommand*{\headauthor}{Porta Mana \etal}
\newcommand*{\reporthead}{}% Report number

@@

The calculations and conclusions presented in the preceding sections depend
on the size $N$ of the larger population. The larger population cannot be
the whole brain, of course. How large should or can $N$ be in our formulae?

The crucial point is our belief distribution for the activity of the sample
\emph{conditional} on the activity of the larger population. It is given by
the hypergeometric distribution~\eqref{eq:conditional_hypergeometric},
reprinted here:
\begin{equation}
  \labelbis{eq:conditional_hypergeometric}
  \yG_{\ya\yA} \defd \pf(\ya \|\yA, n,N)=
    \binom{\yA}{\ya}\binom{N-\yA}{n-\ya}\binom{N}{n}^{-1}
\equiv \binom{n}{\ya}\binom{N-n}{\yA-\ya}\binom{N}{\yA}^{-1}.
\end{equation}
This distribution leads to the equality of factorial
moments~\eqref{eq:equal_m_factorialmoments}, on which our model rests. This
conditional probability, characteristic of \enquote{drawing without
  replacement}
\citep[\eg][\chap~3]{jaynes1994_r2003}, % [\sect~4.8.3]{ross1976_r2010}[\sect~II.6]{feller1950_r1968}
ensues when we believe, with the same degree, that any of the $N$ units
(balls or neurons) could be drawn or sampled. Thus, consider the pool of
all neurons which we believe -- with equal degree for each neuron -- could
have been recorded. Then $N$ is the size of that pool. $N$ therefore depends on
factors such as: the shape, dimensions, and technical specifications of the
recording probe; the inaccuracy in the insertion of the probe, leading for
example to slightly different insertion points or angles; the density of
neurons around the probe. But it also depends on the homogeneity of the
brain region where the recording was made: if we believe that it doesn't
matter whether the probe had been inserted in some other point of the same
brain region, then formula~\eqref{eq:conditional_hypergeometric} is
appropriate. $N$ can therefore only be assessed case by case.

Whether such equal beliefs are justified, or for which sampling procedures
they can be justified, is the fundamental, deep question of sampling
theory, which we cannot discuss here \citep[see \eg\ the discussions and
references
in][]{ericson1969,smith1976}. % {kruskaletal1979c,kruskaletal1980}
It is of course possible to probabilistically assess, case by case, whether
this equality assumption is appropriate. But we must also be aware that
such assessment in turn rests on analogous assumptions at a higher level,
and so on. The probability-calculus, just like the truth-calculus, cannot
yield conclusions without premisses \citep{hailperin2011}[p.~182]{johnson1924}.

Luckily the plots of \fig~\ref{fig:plots_5mom} suggest that the order of
magnitude of $N$ ought to be enough for qualitative inferences. As
discussed in the next section, even if the exact number $N$ were known, the
maximum-entropy distribution ought to be interpreted qualitatively or
semi-quantitatively.

@@

If our uncertainty about $N$ spans several orders of magnitude we can
assign a probability to the possible values of $N$ based on our background
information and the frequency data $\yf$, similarly to the procedure in
\sect~\ref{sec:marginalization} and appendix~\ref{sec:measure_suff}. The
probability for the data $\yf$ conditional on a set of constraints $M$ and
size $N$ is the exponential of the negative relative
entropy~\eqref{eq:relentropy_T}:
\begin{equation}
  \label{eq:prob_data_from_N}
  \pf(\yf \| M, N) = \binom{T}{T\yf}\;\prod_{a}\pxx(\ya \| M, N)^{T\,\yff_{a}}
  \approx \exp\bigl\{  -T\,\sh[\yf; \pxx(\ya \| M, N)]\bigr\}.
\end{equation}
If our pre-data probability for $N$ is $p(N)$, then by Bayes's theorem
\begin{equation}
  \label{eq:prob_N_from_data}
  \pf(N \| M, \yf) \propto
  p(N)\;\exp\bigl\{-T\,\sh[\yf; \pxx(\ya \| M, N)]\bigr\}.
\end{equation}

Here is an illustrative example with our first data set. Suppose our
uncertainty spans slightly more than an order of magnitude, from $N=1\,000$
to $N=20\,000$. Divide this range roughly into thirds of order of
magnitude, considering the values
$N\in \set{1\,000, 2\,000, 5\,000, 10\,000, 20\,000}$. Assuming the set $M$
of first five moments to be sufficient, formula~\eqref{eq:prob_data_from_N}
gives
\begin{equation}
  \label{eq:data1_f_from_N}
  \begin{gathered}
  \pf(\yf \| N=1\,000, M) =  0.00222,\qquad
  \pf(\yf \| N=2\,000, M) =  0.00704,\\
  \pf(\yf \| N=5\,000, M) =  0.0127,\qquad
  \pf(\yf \| N=10\,000, M) =  0.0150,\\
  \pf(\yf \| N=20\,000, M) =  0.0135.
\end{gathered}
\end{equation}
Since our uncertainty regards a scale factor, it can be represented by
equal pre-data probabilities of $1/5$ about these partial orders of
magnitude. Thus from~\eqref{eq:data1_f_from_N} we find
\begin{equation}
  \label{eq:data1_N_from_f}
  \begin{gathered}
  \pf(N=1\,000 \| \yf, M) =  0.044,\qquad
  \pf(N=2\,000 \| \yf, M) =  0.140,\\
  \pf(N=5\,000 \| \yf, M) =  0.251,\qquad
  \pf(N=10\,000 \| \yf, M) =  0.298,\\
  \pf(N=20\,000 \| \yf, M) =  0.267,
\end{gathered}
\end{equation}
which gives a slightly higher probability to $N=10\,000$.

@@

In this way we can make inferences % -- for example about the sufficiency of
% a set of moments, or about the marginal sample distribution --
that take into account our uncertainty about $N$. We must make sure to
avoid circularities, though: for example, we cannot assume a set of moments
$M$ to be sufficient and assess our uncertainty about $N$ conditional on
it, and then use this uncertainty to assess the sufficiency of $M$. But we
can approximately assess the sufficiency of a subset of moments
$M' \subset M$ taking into account the uncertainty of $N$ conditional on
$M$.

@@

The possible uses of the model here presented depend on the assumptions and
approximations which define it: three main assumptions and two main
approximation all in all. These become clear within a full-fledged
probabilistic approach. We summarize them here, leaving a more mathematical
sketch to appendix~\ref{sec:inference_prob}.

The first and most important assumption, typical of this kind of
maximum-entropy application in the neurosciences, can be expressed in two
equivalent ways:
\begin{enumerate*}[(\roman*)]
  % \item the order of the time bins has no relevance;
\item the frequencies of the activity during a recording are
  informationally sufficient for inferences about unrecorded times; \item
  our degree of belief about the sequence of activities is invariant under
  permutations of the time bins.
\end{enumerate*}
Equating time averages and expectations, \eqn~\eqref{eq:final_constraint},
would not make sense without this assumption. Its mathematical consequence
is to relate the probability for the frequency distribution $\yF$ during
the recording to that of the \emph{long-run} frequency distribution for the
activities of the larger population. Note that the maximum-entropy method
in statistical physics is meant to be used not with this assumption, but
with averages across \emph{repetitions} of the same experiment, because its
constraints represent macroscopically \emph{reproducible} quantities and
kinematics, even in non-equilibrium
\citep{jaynes1985b_r1996,%jaynes1957,%tikochinskyetal1984,
grandy1988,%berryetal1988,
deroecketal2005_r2006%,touchette2009
}[for experiments \eg][]{guntonetal1983}.

The second assumption, typical of maximum-entropy in general, concerns the
form of the probability distribution for the long-run frequencies ensuing
from the first assumption. It is taken to be either an \enquote{entropic
  prior} \citep{skillingetal1984,rodriguez1991,neumann2007}, which takes
into account the multiplicity of long-run frequencies, or a prior for a
model with sufficient statistics, which expresses the sufficiency of the
moments under study \citep{portamana2017}. This assumption is crucial for
deriving the maximum-entropy based on the Shannon entropy. A different
density leads to alternative maximum-entropy methods, for example based on
the Burg entropy \citep{burg1975} if a Dirichlet density is used
\citep{jaynes1986d_r1996}. The entropic prior depends on a reference
distribution and on a coefficient $L$ which determines the sharpness of the
prior's peak. The computations of the preceding sections were insensitive
to the reference distribution in their significant digits. We used three
biologically reasonable alternatives: a flat distribution, one linearly
decreasing with $\yA$, and a normal one with a broad peak at low
activities.

The third assumption is specific to our application: all subsets of $\yA$
neurons in the larger population are equally likely to give rise to total
activity $\yA$, at every time bin. A mathematically equivalent assumption
is that $n$ neurons are sampled anew at each time bin. This assumption
allows us to factorize the joint probability of the two sequences of
activities and apply the sampling
relation~\eqref{eq:conditional_hypergeometric} of
appendix~\ref{sec:derivation_maxent_fullnet} at each bin.

Finally, the maximum-entropy method is typically equivalent to
approximating the number $T$ of time bins, the prior coefficient $L$, and
the ratio $T/L$ to infinity. These approximations lead to the
identification of the four distributions listed at the end of
\sect~\ref{sec:model}: the recorded frequency distribution becomes equal to
the long-run one, and also to the probability distribution for the activity
at recorded and new time bins. In our application it also leads to the
equality of the hypergeometric
distribution~\eqref{eq:conditional_hypergeometric} and the conditional
frequencies of the sample activity $\ya$ given $\yA$. The goodness of this
approximation decreases as the ratio between the number of possible joint
states and the number of bins -- roughly $nN/T$ -- increases, and also as
observed quantities approach their mathematical bounds, such as some
frequencies $\yff_{\ya}=0$. Without these approximations it is still
possible to interpret the maximum-entropy
distribution~\eqref{eq:app_maxent_pop_first} as the mode of the probability
distribution for the recorded frequencies $\yF$, that is, the most probable
recorded frequency.

@@

The first and third assumptions above are the least realistic, but they
make sense as reference hypotheses for judging the informational relevance
of correlations of some order, as in \sect~\ref{sec:marginalization}, to be
compared with hypotheses about more realistic time dependences and
clustering. %They may turn out to approximate well more realistic assumptions.

The $T= \infty$ approximation can break our model for specific
distributions of sample frequencies, if too many moments are used. For
example take $n=2$ with frequency distribution $\yf=(0.5,0,0.5)$ for the
activities $\ya=0,1,2$; that is, the sample is silent half of the time and
completely active half of the time. The first two normalized factorial
moments have both value $0.5$, and together they uniquely determine $\yf$.
If we now take $N=3$, it can be proved that
\eqn~\eqref{eq:app_maxent_pop_first}, constrained on those two moments, has
no solution. The reason is clear from a sampling perspective: if there is
zero probability of sampling one active neuron, then the larger population
cannot have any active neurons, and there must be zero probability of
sampling two active ones. But this reasoning holds only for each single
time bin, not for the averages over time bins. The $T=\infty$ approximation
identifies the two instead. Outside this approximation there are no
contradictions because the conditional frequencies and conditional
probabilities are kept distinct.

The two infinity approximations are at the limit of their validity in our
examples. For this reason we interpret the distributions as \emph{most
  probable frequency distributions}, as mentioned at the end of
\sect~\ref{sec:model}. The space of possible frequency distributions for
the larger population has $N$ dimensions, however, and probability
distributions in high dimensions have counter-intuitive properties. For
example, the mode or mean can have \emph{atypical} features when compared
with the majority of other probable points of the probability space. The
question then is \emph{which features of the maximum-entropy frequency
  distribution are typical of the majority of plausible frequency
  distributions?} Preliminary studies using a full probability calculation
(to be published separately) show that several semi-quantitative features
of the maximum-entropy distribution are indeed typical. For the first data
set, for example, most frequency distributions have two high-frequency
regimes of activity, as in \fig~\ref{fig:plots_5mom_a}, with roughly the
same heights and almost the same locations, within a few percentages of
population-averaged activity. Each of high-frequency regimes, though, may
consist of two or three very close modes rather than one. The same
preliminary studies also indicate that these typical features remain with
prior choices in the second assumption above, for example using a Dirichlet
prior instead of an entropic one.

This means that the maximum-entropy approximation presented here can be
favourably used to get an idea of what a fully probabilistic calculation
could yield, but with a far lower computational cost. For example, for
$N=1\,000$ the maximum-entropy distribution can be calculated in ten
minutes on a laptop, whereas the full probability calculation takes several
days on a high-performance computing cluster.

>>>>

\pdfoutput=1
%% Author: PGL  Porta Mana
%% Created: 2015-05-01T20:53:34+0200
%% Last-Updated: 2019-12-22T15:04:32+0100
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newif\ifarxiv
\arxivfalse
\ifarxiv\pdfmapfile{+classico.map}\fi
\newif\ifafour
\afourfalse% true = A4, false = A5
\newif\iftypodisclaim % typographical disclaim on the side
\typodisclaimtrue
\newcommand*{\memfontfamily}{zplx}
\newcommand*{\memfontpack}{newpxtext}
\documentclass[\ifafour a4paper,12pt,\else a5paper,10pt,\fi%extrafontsizes,%
onecolumn,oneside,article,%french,italian,german,swedish,latin,
british%
]{memoir}
\newcommand*{\updated}{\today}
\newcommand*{\firstdraft}{4 November 2015}
\newcommand*{\firstpublished}{***}
\newcommand*{\propertitle}{Inferring the total activity of a large neuronal population\\ from a small sample}
\newcommand*{\pdftitle}{\propertitle}
\newcommand*{\headtitle}{Inferences about large neuronal populations}
\newcommand*{\pdfauthor}{P.G.L.  Porta Mana, Y. Roudi, V. Rostami, E. Torre}
\newcommand*{\headauthor}{Porta Mana \etal}
\newcommand*{\reporthead}{}% Report number

@@

The calculations presented in the preceding sections depend on the size $N$
of the larger population, which of course cannot be the whole brain. How
large should or can $N$ be in our formulae?

The crucial point is our belief distribution for the activity of the sample
\emph{conditional} on the activity of the larger population. It is given by
the hypergeometric distribution~\eqref{eq:conditional_hypergeometric},
reprinted here:
\begin{equation}
  \labelbis{eq:conditional_hypergeometric}
  \yG_{\ya\yA} \defd \pf(\ya \|\yA, n,N)=
    \binom{\yA}{\ya}\binom{N-\yA}{n-\ya}\binom{N}{n}^{-1}
\equiv \binom{n}{\ya}\binom{N-n}{\yA-\ya}\binom{N}{\yA}^{-1}.
\end{equation}
This distribution leads to the equality of factorial
moments~\eqref{eq:equal_m_factorialmoments}, on which our model rests. This
conditional probability, characteristic of \enquote{drawing without
  replacement}
\citep[\eg][\chap~3]{jaynes1994_r2003}, % [\sect~4.8.3]{ross1976_r2010}[\sect~II.6]{feller1950_r1968}
ensues when we believe, with the same degree, that any of the $N$ units
(balls or neurons) could be drawn or sampled. Thus, consider the pool of
all neurons which we believe -- with equal degree of belief for each neuron
-- could have been recorded. Then $N$ is the size of that pool. $N$
therefore depends on factors such as: the shape, dimensions, and technical
specifications of the recording probe; the inaccuracy in the insertion of
the probe, leading for example to slightly different insertion points or
angles; the density of neurons around the probe. But it also depends on the
homogeneity of the brain region where the recording is made: if we believe
that it does not matter whether the probe had been inserted in some other
point of the same brain region, then
formula~\eqref{eq:conditional_hypergeometric} is appropriate. $N$ can
therefore only be assessed case by case.

Whether such symmetric beliefs are justified, or for which sampling procedures
they can be justified, is the fundamental, deep question of sampling
theory, which we cannot discuss here \citep[see \eg\ the discussions and
references
in][]{ericson1969,smith1976}. % {kruskaletal1979c,kruskaletal1980}
It is of course possible to probabilistically assess, case by case, whether
this symmetry assumption is appropriate. But we must also be aware that
such assessment in turn rests on analogous symmetry assumptions at a higher level,
and so on. The probability-calculus, just like the truth-calculus, cannot
yield conclusions without premisses \citep{hailperin2011}[p.~182]{johnson1924}.

The plots of \fig~\ref{fig:plots_5mom} nevertheless suggest that the order
of magnitude of $N$ ought to be enough for qualitative inferences. As
discussed in the next section, even if the exact number $N$ were known, the
maximum-entropy distribution ought to be interpreted qualitatively or
semi-quantitatively.

@@

If our uncertainty about $N$ spans several orders of magnitude we can
assign a probability to the possible values of $N$ based on our background
information and the frequency data $\yf$, similarly to the procedure in
\sect~\ref{sec:marginalization} and appendix~\ref{sec:measure_suff}. The
probability for the data $\yf$ conditional on a set of constraints $M$ and
size $N$ is the exponential of the negative relative
entropy~\eqref{eq:relentropy_T}:
\begin{equation}
  \label{eq:prob_data_from_N}
  \pf(\yf \| M, N) \approx \binom{T}{T\yf}\;\prod_{a}\pxx(\ya \| M, N)^{T\,\yff_{a}}
  \approx \exp\bigl\{  -T\,\sh[\yf; \pxx(\ya \| M, N)]\bigr\}.
\end{equation}
If our pre-data probability for $N$ is $p(N)$, then by Bayes's theorem
\begin{equation}
  \label{eq:prob_N_from_data}
  \pf(N \| M, \yf) \propto
  p(N)\;\exp\bigl\{-T\,\sh[\yf; \pxx(\ya \| M, N)]\bigr\}.
\end{equation}

Here is an illustrative example with our first data set. Suppose our
uncertainty spans slightly more than one order of magnitude, from $N=1\,000$
to $N=20\,000$. Divide this range roughly into thirds of order of
magnitude, considering the values
$N\in \set{1\,000, 2\,000, 5\,000, 10\,000, 20\,000}$. Assuming the set $M$
of first five moments to be sufficient, formula~\eqref{eq:prob_data_from_N}
gives
\begin{equation}
  \label{eq:data1_f_from_N}
  \begin{gathered}
  \pf(\yf \| N=1\,000, M) =  0.00222,\qquad
  \pf(\yf \| N=2\,000, M) =  0.00704,\\
  \pf(\yf \| N=5\,000, M) =  0.0127,\qquad
  \pf(\yf \| N=10\,000, M) =  0.0150,\\
  \pf(\yf \| N=20\,000, M) =  0.0135.
\end{gathered}
\end{equation}
Since our uncertainty regards a scale factor, it can be represented by
equal pre-data probabilities of $1/5$ about these partial orders of
magnitude. Thus from~\eqref{eq:data1_f_from_N} we find
\begin{equation}
  \label{eq:data1_N_from_f}
  \begin{gathered}
  \pf(N=1\,000 \| \yf, M) =  0.044,\qquad
  \pf(N=2\,000 \| \yf, M) =  0.140,\\
  \pf(N=5\,000 \| \yf, M) =  0.251,\qquad
  \pf(N=10\,000 \| \yf, M) =  0.298,\\
  \pf(N=20\,000 \| \yf, M) =  0.267,
\end{gathered}
\end{equation}
which gives a slightly higher probability to $N=10\,000$.

@@

In this way we can make inferences % -- for example about the sufficiency of
% a set of moments, or about the marginal sample distribution --
that take into account our uncertainty about $N$. We must make sure to
avoid circularities: for example, we cannot assume a set of moments
$M$ to be sufficient and assess our uncertainty about $N$ conditional on
it, and then use this uncertainty to assess the sufficiency of $M$. But we
can approximately assess the sufficiency of a subset of moments
$M' \subset M$ taking into account the uncertainty of $N$ conditional on
$M$.

@@

The possible uses of the model here presented depend on the assumptions and
approximations which define it: three main assumptions and two main
approximations all in all. They become clear within a full-fledged
probabilistic approach. We summarize them here, leaving a more mathematical
sketch to appendix~\ref{sec:inference_prob}.

The first and most important assumption, typical of this kind of
maximum-entropy application in the neurosciences, can be expressed in two
equivalent ways:
\begin{enumerate*}[(\roman*)]
  % \item the order of the time bins has no relevance;
\item the frequencies of the activity during a recording are
  informationally sufficient for inferences about unrecorded times; \item
  our degree of belief about the sequence of activities is invariant under
  permutations of the time bins.
\end{enumerate*}
Equating time averages and expectations, \eqn~\eqref{eq:final_constraint},
would not make sense without this assumption. Its mathematical consequence
is to relate the probability for the frequency distribution $\yF$ for the
activity of the larger population during the recording, to the probability
for the \emph{long-run} frequency distribution. In statistical physics, as
a comparison, the maximum-entropy method is instead used with the
assumption of permutation-invariance across \emph{repetitions} of the same
experiment, because its constraints represent macroscopically
\emph{reproducible} quantities and kinematics, even in non-equilibrium
\citep{jaynes1985b_r1996,%jaynes1957,%tikochinskyetal1984,
  grandy1988,%berryetal1988,
  deroecketal2005_r2006%,touchette2009
}[for experiments \eg][]{guntonetal1983}. The kinetic-Ising models
mentioned in \sect~\ref{sec:intro} are closer to this latter approach.

The second assumption, typical of maximum-entropy in general, concerns the
probability distribution for the long-run frequencies ensuing from the
first assumption \citep[and references therein]{portamana2017}. It is taken
to be either an \enquote{entropic prior}
\citep{skillingetal1984,rodriguez1991,neumann2007}, which takes into
account the multiplicity of long-run frequencies, or a prior for a model
with sufficient statistics, which expresses the sufficiency of the moments
under study, according to the theorem discussed in
\sect~\ref{sec:marginalization}. This assumption is crucial for the
maximum-entropy method based on the Shannon entropy. Different prior
densities lead to alternative maximum-entropy methods; for example one
based on the Burg entropy \citep{burg1975} if a Dirichlet density is used
\citep{jaynes1986d_r1996}. The entropic prior depends on a reference
distribution and on a coefficient $L$ which determines the sharpness of the
prior's peak. The computations of \sects~\ref{sec:application}
and~\ref{sec:marginalization} were insensitive to the choice of the
reference distribution in their significant digits. We used three
biologically reasonable alternatives: a flat distribution, one linearly
decreasing with $\yA$, and a normal one with a broad peak at low
activities.

The third assumption is specific to our application: all subsets of $\yA$
neurons in the larger population are equally likely to give rise to total
activity $\yA$, at every time bin. A mathematically equivalent assumption
is that $n$ neurons are sampled anew at each time bin. This assumption
allows us to factorize the joint probability of the sequence of activities
and apply the sampling relation~\eqref{eq:conditional_hypergeometric} of
appendix~\ref{sec:derivation_maxent_fullnet} at each bin.

Finally, the maximum-entropy method is typically equivalent to
approximating the number $T$ of time bins, the prior coefficient $L$, and
the ratio $T/L$ to infinity. These approximations lead to the
identification of the four distributions listed at the end of
\sect~\ref{sec:model}: the recorded frequency distribution becomes equal to
the long-run one, and also to the probability distribution for the activity
at recorded and new time bins. In our application it also leads to the
equality of the hypergeometric
distribution~\eqref{eq:conditional_hypergeometric} and the conditional
frequencies of the sample activity $\ya$ given $\yA$. The goodness of this
approximation decreases as the ratio -- roughly $nN/T$ -- between the
number of possible joint states and the number of bins increases, and also
as observed quantities approach their mathematical bounds, for example some
frequencies $\yff_{\ya}=0$. Without these approximations it is still
possible to interpret the maximum-entropy
distribution~\eqref{eq:app_maxent_pop_first} as the mode of the probability
distribution for the recorded frequencies $\yF$, that is, the most probable
recorded frequency.

@@

The first and third assumptions above are the least realistic, but they
make sense as reference hypotheses for judging the informational relevance
of correlations of some order, as discussed in
\sect~\ref{sec:marginalization}, to be compared with hypotheses about more
realistic time dependences and
clustering. %They may turn out to approximate well more realistic assumptions.

The $T= \infty$ approximation can break our model for specific
distributions of sample frequencies, if too many moments are used. For
example, take $n=2$ with frequency distribution $\yf=(0.5,0,0.5)$ for the
activities $\ya=0,1,2$; that is, the sample is silent half of the time and
completely active half of the time. The first two normalized factorial
moments have both value $0.5$, and together they uniquely determine $\yf$.
If we now take $N=3$, it can be proved that
\eqn~\eqref{eq:app_maxent_pop_first}, constrained on those two moments, has
no solution. The reason is clear from a sampling perspective: if there is
zero probability of sampling one active neuron, then the larger population
cannot have any active neurons at all, and a fortiori there must be zero
probability of sampling two active ones. But this reasoning holds only for
each single time bin, not for the averages over time bins. The $T=\infty$
approximation identifies the two instead. Outside of this approximation
there are no contradictions because the conditional frequencies and
conditional probabilities are kept distinct.

The two infinity approximations are at the boundary of their validity in
our examples. For this reason we interpret the maximum-entropy distribution
as the \emph{most probable frequency distributions}, as mentioned at the
end of \sect~\ref{sec:model}. An important question, then, is \emph{which
  features of the maximum-entropy frequency distribution are typical of the
  majority of plausible frequency distributions?} Especially for features
such as number, position, height of modes, or tail behaviour. Preliminary
studies using a full probability calculation (to be published separately)
show that several semi-quantitative features of the maximum-entropy
distribution are indeed typical. For the first data set, for example, most
frequency distributions have two high-frequency regimes of activity, as in
\fig~\ref{fig:plots_5mom_a}, with roughly the same heights and almost the
same locations, within a few percentages of population-averaged activity.
Each of the two high-frequency regimes, though, may consist of two or three
very close modes rather than one. The same preliminary studies also
indicate that these typical features are robust for different prior choices
in the second assumption above, for example using a Dirichlet prior instead
of an entropic one.

This means that the maximum-entropy approximation presented here can be
favourably used to get an idea of what a fully probabilistic calculation
could yield -- but with a far lower computational cost than the latter. For
example, for $N=1\,000$ the maximum-entropy distribution can be calculated
in ten minutes on a laptop, whereas the full probability calculation takes
several days on a high-performance computing cluster.
